Injury Outcomes

### 2023-08-21 injury_tables.Rmd
updated comments

smodes <- read_csv('data/global/modes/standardized_modes.csv')

@@

# WARNING: Script assumes that the Baseline scenario corresponds to the known injury fatality counts in the city, i.e it assumes that the reference scenario is the Baseline scenario

@@

## - html file containing the following information:

#   - fatality counts by strike and casualty mode for each scenario and city (incl lower and upper confidence interval limits were applicable), one table for each city, scenario and mean, upper or lower confidence interval values

#   - fatality counts giving lower, mean and upper bounds (where they exist) for each casualty mode by strike mode for each scenario and city, one table for each casualty mode, scenario and city

#   - tables showing the fatalities per billion km for each city, scenario and mean, upper and lower confidence interval values (where applicable).

@@

#   - tables showing the fatalities per 100 million hours for each city, scenario and mean, upper and lower confidence interval values (where applicable).

## - output .csv files containing the following information:

@@

## The script performs the following steps, assuming that ITHIM-GLOBAL module has 
# been run and the ithim_objects object has been saved in "results/multi_city/io.rds":

# - read in ithim_objects and extract relevant information such as city and scenario names 
# - define various functions to summarise and print data and to re-name scenarios, also
#   define function to collapse the age categories into two categories (15-49 and 50-69)
# - calculate the health burden (yll and deaths) and the demographic information for those
#   two age categories (15-49 and 50-69)

# - loop through cities:
#   - loop through the mean injury values but also the upper and lower confidence boundary values if given:
#     - loop through scenarios incl baseline:
#       - create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
#         of nov or whw exists, only create dataframe with existing data), add row and column sums
#       - delete casualty and strike modes that do not exist in list of standardized modes
#       - print dataframe to html file
#       - create one list containing the predicted injury counts for each city, each scenario and for each scenario
#         both the mean and the upper and lower confidence interval limits are given if they exist
#       - create a dataframe with the original fatality counts for each casualty and strike mode pair
#         (adjusted by weight and injury reporting rate)
#       - create dataframe containing the mode distances for all scenarios including the predicted baseline
#       - create dataframes containing the population counts, mode speeds and durations travelled

# - create dataframe that contains the fatality counts for each city, scenario and casualty and strike mode pair.
#   It also includes the observed fatality counts and the active travel fatality counts (the sum of cycle and 
#   walk casualty mode fatalities) - ('results/multi_city/inj/inj_counts.csv') 
#   - print individual table for each city, scenario and mean, upper and lower confidence value fatality counts 
#     (where applicable) by casualty and strike mode to html file

@@

# - For each city and scenario give mean (and upper and lower confidence boundaries where required) fatalities per 
#   billion km travelled for each casualty and strike mode pair. Create html tables.

@@

# - Create dataframe containing all mode distances for each city and scenario ('distances.csv')

@@

# - Export all injury datasets into a single excel sheet (inj_data.xlsx)

@@

# Remove extra bits of info - irrelevant outputs for injury related outputs
# This makes it the io object lighter
for (city in cities){
  io[[city]][["trip_scen_sets"]] <- NULL
  io[[city]][["inj_distances"]] <- NULL
  io[[city]][["synth_pop"]] <- NULL

  io[[city]]$outcomes$mmets <- io[[city]]$outcomes$pm_conc_pp <- NULL
}

@@

# function to round and sum the data and then to print it 
# the html output document
sum_and_round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  data <- lapply(data,function(x)rbind(x,Total=colSums(x)))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

@@

```{r preprocessing}
# start the processing

# function to collapse the given age categories into only two categories
# '15-49' and '50-69'
collapse_ages <- function(data, ages = c('15-49', '50-69'), age_lab = 'age_cat'){
  target_min_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][1]))
  target_max_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][2]))
  min_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][1]))
  max_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][2]))
  genders <- unique(data$sex)
  if (ncol(data) > 3) {
    reformatted <- do.call(rbind,lapply(1:length(ages),
                                        function(x) t(sapply(genders, function(y) 
                                          if(ncol(data)>3)     colSums(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                                             else sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                        ))
    ))
    data.frame(age_cat=rep(ages,each=2),sex=rep(genders,2),reformatted,stringsAsFactors=F)
  }else{
    reformatted <- do.call(c,lapply(1:length(ages),
                                    function(x) sapply(genders,
                                                       function(y) 
                                                         sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                    )
    ))
    data.frame(age=rep(ages,each=2),sex=rep(genders,2),population=as.numeric(reformatted),stringsAsFactors=F)
  }
}

scen_prop <- io$scen_prop
io <- io[-2]

# re-format some age and injury names
for (i in 1:length(io)) {
  io[[i]]$demographic$age[io[[i]]$demographic$age == '5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$deaths$age_cat[io[[i]]$outcomes$pathway_hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$ylls$age_cat[io[[i]]$outcomes$pathway_hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$deaths <-      io[[i]]$outcomes$pathway_hb$deaths[,!sapply(names(io[[i]]$outcomes$pathway_hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$ylls <- io[[i]]$outcomes$pathway_hb$ylls[,!sapply(names(io[[i]]$outcomes$pathway_hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
}

# calculate the demographics for the two age categories '15-49' and '50-69'
for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')  # only keep two age categories

@@

```{r whw, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
# create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
# of nov or whw exists, only create dataframe with existing data), add row and column sums
# create distance, speed and duration data frames

@@

    # loop through scenarios incl baseline
    for (cs in names(io[[city]]$outcomes$whw)){ 
      # cs <- 'baseline'

@@

      if (length(city_df[[cs]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
        all_whw[[cs]][[city]] <- data.table::rbindlist(city_df[[cs]][[city]]) # combine mean, lb and ub counts into one table for each scenario
    }
  }

@@

    # recalculate value based on weight and injury reporting rate
    injury_data_count_nov$value <- injury_data_count_nov$value / injury_data_count_nov$weight / injury_data_count_nov$injury_reporting_rate
    injury_data_count_nov <- injury_data_count_nov[c('str_mode', 'value', 'cas_mode')] # re order columns and drop column weight

@@

    city_baseline_count <- rbind(injury_data_count2, injury_data_count_nov2)
  } else {
    city_baseline_count <- injury_data_count2
  }

@@

  col_split <- stringr::str_split(inj_counts$variable, "_", simplify = TRUE, n = 2) # create one city and one strike mode column

@@

  #qualified_scen_name <- 'Public Transport'
  scen <- 'Scenario'

@@

# add original counts and rename Baseline to Baseline predicted
#inj_counts_list$scenario <- ifelse(inj_counts_list$scenario == 'Baseline', 'Baseline_predicted',inj_counts_list$scenario)

@@

# Remove total from whw long form, and write as csv
readr::write_csv(inj_counts_list, paste0('results/multi_city/inj/inj_counts.csv')) # table containing all the fatality counts
```

@@

```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
### Tables for each casualty mode, giving fatalities by strike mode -> print to html file

@@

    index <- index + 1
    cat('\n')
    cat('#### Casualty mode: ', stringr::str_to_title(cm), '\n')
    #print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))

    ft <- flextable(temp1)
    ft <- add_header_row(ft, values = c(' ', 'strike mode'), colwidths = c(3, ncol(temp1) - 3))
    ft <- merge_v(ft, j = 1:3, part = 'body')

@@

@@

      # add city's population
      pop <- sum(io[[city]]$demographic$population)
      # Compute distance travelled per capita
      td6 <- td5
      td6[[paste0(var,"_risk")]] <- td4[2]
      td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop # calculate fatalities per capita
      names(td6)[3] <- paste(names(td6)[2], names(td6)[3], sep = "_")
      dd <- full_join(dd, td6, by = 'stage_mode') 
    }

@@

# add distance information (note true_distances does not contain distance info for active_travel)
rd <- left_join(rd, true_distances, by = c('mode', 'city','scenario'))

@@

# add total counts (excluding active travel)

@@

    # calculate total per scenario (not including active travel)
    #el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))

@@

# add country and continent information
# injury_risks_per_100k <- left_join(injury_risks_per_100k, location, by = 'city')

readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/inj/injury_risks_per_100k_pop.csv')
```

@@

        # td3: number of fatalities for whw only
        td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")

@@

    td <- el # %>% dplyr::select(-contains('count'))
    # add city name to column names
    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
    # add 'measure' columns
    td$measure <- ifelse(is.na(pref), "mean", pref)

@@

```{r}
# Export injury datasets in a single excel file
path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"
 writexl::write_xlsx(list(inj_counts = inj_counts_list, #%>% filter(str_mode != 'Total'),
                          injury_risks_billion_kms = rd,
                          injury_risks_100k_people  = injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'),
                          injury_risks_100mil_h = rd_hours,
                          true_distances = true_distances_df),
                     path = path_inj_wb)
```

>>>>

@@

# WARNING: Script assumes that the Baseline scenario corresponds to the known injury fatality counts in the city

@@

### - html file containing the following information:

#   - fatality counts by strike and casualty mode for each scenario and city (incl lower and upper confidence interval limits were applicable); one table for each city, scenario and mean, upper or lower confidence interval values

#   - fatality counts giving lower, mean and upper bounds (where they exist) for each casualty mode by strike mode for each scenario and city; one table for each casualty mode, scenario and city

#   - tables showing the fatalities per billion km travelled for each city, scenario and mean, upper and lower confidence interval values (where applicable).

@@

#   - tables showing the fatalities per 100 million hours travelled for each city, scenario and mean, upper and lower confidence interval values (where applicable).

### - output .csv files containing the following information:

@@

## The script performs the following steps, assuming that ITHIM-GLOBAL has been run and the ithim_objects object has been saved in "results/multi_city/io.rds":

# - read in ithim_objects and extract relevant information such as city and scenario names 
# - define various functions to summarise and print data and to re-name scenarios

# - loop through cities:
#   - loop through the mean injury values but also the upper and lower confidence boundary values if given:
#     - loop through scenarios incl baseline (original counts) and baseline predicted:
#       - create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
#         of nov or whw exists, only create dataframe with existing data), add row and column sums
#       - delete casualty and strike modes that do not exist in list of standardized modes
#       - print dataframe to html file
#       - create one list containing the predicted injury counts for each city, each scenario and for each scenario
#         both the mean and the upper and lower confidence interval limits are given if they exist
#       - create a dataframe with the original fatality counts for each casualty and strike mode pair
#         (adjusted by weight and injury reporting rate)
#       - create dataframe containing the mode distances for all scenarios including the predicted baseline
#       - create dataframes containing the population counts, mode speeds and durations travelled

# - create dataframe that contains the fatality counts for each city, scenario and casualty and strike mode pair.
#   It also includes the observed fatality counts and the active travel fatality counts (the sum of cycle and 
#   walk casualty mode fatalities) ('results/multi_city/inj/inj_counts.csv')

@@

# - For each city and scenario give mean (and upper and lower confidence boundaries where required) fatalities per 
#   billion km travelled for each casualty mode. Create html tables.

@@

# - Create dataframe containing all daily population mode distances for each city and scenario ('distances.csv')

@@

# - Export all injury datasets into a single excel file (inj_data.xlsx)

# read in table detailing the standardized modes
smodes <- read_csv('data/global/modes/standardized_modes.csv')

@@

# Remove extra bits of info - irrelevant outputs for injury related outputs
# This makes the io object lighter
for (city in cities){
  io[[city]][["trip_scen_sets"]] <- NULL
  io[[city]][["inj_distances"]] <- NULL
  io[[city]][["synth_pop"]] <- NULL
  io[[city]]$outcomes$mmets <- io[[city]]$outcomes$pm_conc_pp <- NULL
}

@@

# function to round and sum the data and then to print it to
# the html output document
sum_and_round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  data <- lapply(data,function(x)rbind(x,Total=colSums(x)))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

@@

@@

```{r whw, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
# create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
# of nov or whw exists, only create dataframe with existing data), add row and column sums
# create distance, speed and duration dataframes

@@

    # loop through scenarios incl baseline predicted
    for (cs in names(io[[city]]$outcomes$whw)){ 
      # cs <- 'baseline'

@@

      if (length(city_df[[cs]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
        # combine mean, lb and ub counts into one table for each scenario
        all_whw[[cs]][[city]] <- data.table::rbindlist(city_df[[cs]][[city]]) 
    }
  }

@@

    # recalculate value based on weight and injury reporting rate
    injury_data_count_nov$value <- injury_data_count_nov$value / injury_data_count_nov$weight / injury_data_count_nov$injury_reporting_rate
    # re order columns and drop column weight
    injury_data_count_nov <- injury_data_count_nov[c('str_mode', 'value', 'cas_mode')]

@@

    # combine whw and nove data
    city_baseline_count <- rbind(injury_data_count2, injury_data_count_nov2)
  } else {
    city_baseline_count <- injury_data_count2
  }

@@

  # create one city and one strike mode column
  col_split <- stringr::str_split(inj_counts$variable, "_", simplify = TRUE, n = 2)

@@

  scen <- 'Scenario'

  # rename scenarios
  qualified_scen_name <- get_qualified_scen_name(cs)

@@

@@

# write to csv
readr::write_csv(inj_counts_list, paste0('results/multi_city/inj/inj_counts.csv')) # table containing all the fatality counts
```

@@

```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}

### Tables for each casualty mode, giving fatalities by strike mode -> print to html file

@@

    index <- index + 1
    cat('\n')
    cat('#### Casualty mode: ', stringr::str_to_title(cm), '\n')

    # updated formating 
    ft <- flextable(temp1)
    ft <- add_header_row(ft, values = c(' ', 'strike mode'), colwidths = c(3, ncol(temp1) - 3))
    ft <- merge_v(ft, j = 1:3, part = 'body')

@@

      # get updated scenario names
      var <- get_qualified_scen_name(cs)

@@

      # add city's population
      pop <- sum(io[[city]]$demographic$population)

       # Compute distance travelled per capita
      td6 <- td5
      td6[[paste0(var,"_risk")]] <- td4[2]
      td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop # calculate fatalities per capita
      names(td6)[3] <- paste(names(td6)[2], names(td6)[3], sep = "_")
      dd <- full_join(dd, td6, by = 'stage_mode') 
    }

@@

# add distance information 
rd <- left_join(rd, true_distances, by = c('mode', 'city','scenario'))

@@

@@

@@

# write to csv
readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/inj/injury_risks_per_100k_pop.csv')
```

@@

        # td3: number of fatalities for nov only
        td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")

@@

    td <- el # %>% dplyr::select(-contains('count'))

    # add city name to column names
    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
    # add 'measure' columns
    td$measure <- ifelse(is.na(pref), "mean", pref)

@@

```{r}
# Export injury datasets as a single excel file
path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"
 writexl::write_xlsx(list(inj_counts = inj_counts_list, #%>% filter(str_mode != 'Total'),
                          injury_risks_billion_kms = rd,
                          injury_risks_100k_people  = injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'),
                          injury_risks_100mil_h = rd_hours,
                          true_distances = true_distances_df),
                     path = path_inj_wb)
```
